---
title: "Homework 3"
subtitle: "Diagnostics and GLM"
author: "Brian Motherway"
output: pdf_document
---

```{r setup, message=FALSE, warning=FALSE, error=FALSE}
library(ISLR2)
library(car)
library(leaps)
library(msm)
library(tidyverse)
library(stringr)
library(GLMsData)
library(statmod)
library(MASS)
```

# Question 1
```{r}
setwd("C:/Users/mothe/OneDrive/Documents/CMU/Mini 2/Statistical Foundations of Business Analytics/Data")
real_estate <- read_csv("real-estate-valuation-data-set.csv")

multi_reg <- lm(unit_price ~ convenience_stores + log(distance), data = real_estate)
summary(multi_reg)
```

# Question 2
```{r}
std_residuals <- rstandard(multi_reg)
mean(std_residuals)
var(std_residuals)

scatter.smooth(
  std_residuals ~ real_estate$convenience_stores,
  col = "grey",
  las = 1,
  ylab = "Standardized residuals",
  xlab = "Height (inches)"
)

res_conv_plot <- ggplot(data = real_estate, aes(x = real_estate$convenience_stores, y = std_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_conv_plot
```
Generally the residuals are fairly consistent with a average standardized residual of about 0. However, there is one data point at height 3 that has and standardized residual of 8, I would want to look into this data point further with outlier analysis.

# Question 3
```{r}
res_logdis_plot <- ggplot(data = real_estate, aes(x = log(real_estate$distance), y = std_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_logdis_plot
```
Yes, the standardized residuals for value of log(distance) have a trend below zero. I would want to look into this further, we may be over estimating the value for those points.

# Question 4
```{r}
multi_reg_fitted <- fitted(multi_reg)
res_funitprice_plot <- ggplot(data = real_estate, aes(x = multi_reg_fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_funitprice_plot 
```
Yes, the standardized residuals for a unit price of greater than 50 is not averaging zero. We may over estimating these points.

#Question 5
```{r}
qqnorm(std_residuals, las = 1, pch = 19)
qqline(std_residuals)
```
It looks like the model might be assuming tails that are slightly too heavy for the model.

#Question 6
```{r}
box_cox <- boxcox(unit_price ~ convenience_stores + log(distance), data = real_estate)
lambda <- box_cox$x[which.max(box_cox$y)] 
lambda
```
#Question 7 
```{r}
bc_mlr <- lm(((unit_price^lambda - 1)/ lambda)~ convenience_stores + log(distance),  data = real_estate)
summary(bc_mlr)
hist(multi_reg$residuals) #Original residuals
hist(bc_mlr$residuals) # Much more Normal-looking residuals.
```
#Question 8
##8.2
```{r}
bcstd_residuals <- rstandard(bc_mlr)
mean(bcstd_residuals)
var(bcstd_residuals)

scatter.smooth(
  bcstd_residuals ~ real_estate$convenience_stores,
  col = "grey",
  las = 1,
  ylab = "Standardized residuals",
  xlab = "Height (inches)"
)

bcres_conv_plot <- ggplot(data = real_estate, aes(x = real_estate$convenience_stores, y = bcstd_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_conv_plot
bcres_conv_plot
```
##8.3
```{r}
bcres_logdis_plot <- ggplot(data = real_estate, aes(x = log(real_estate$distance), y = bcstd_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_logdis_plot
bcres_logdis_plot
```
There is no material difference in the residuals for the boxcox plot.

##8.4
```{r}
bc_mlr_fitted <- fitted(bc_mlr)
bcres_funitprice_plot <- ggplot(data = real_estate, aes(x = bc_mlr_fitted, y = bcstd_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_funitprice_plot 
bcres_funitprice_plot
```
There is no material difference in the residuals for the boxcox plot.

##8.5
```{r}
qqnorm(std_residuals, las = 1, pch = 19)
qqline(std_residuals)
qqnorm(bcstd_residuals, las = 1, pch = 19)
qqline(bcstd_residuals)
```
There is no material difference in the normality for the boxcox plot.

# Question 9
```{r}
bc_mlr_im <- influence.measures(bc_mlr)
summary(bc_mlr_im)
```
No, no observation for DFBETA were flagged as influential

#Question 10
```{r}
?germ
data(germ)
summary(germ)

glmmodel <- glm(
  Germ / Total ~ Extract + Seeds,
  weights = Total,
  family = "binomial",
  data = germ
)
summary(glmmodel)
```
# Question 11

##11.1
```{r}
levels(germ$Extract)
levels(germ$Seeds)
```
The baseline categories are Bean for Extract and OA73 for Seed

##11.2
```{r}
exp(glmmodel$coefficients[1])
```
The odds of germination for the baseline combination is .4965605

##11.3
```{r}
exp(glmmodel$coefficients[2])
```
The odds are 2.900113 times greater.

##11.4
```{r}
exp(glmmodel$coefficients[3])
```
The odds are 1.310555 times greater.

##11.5
```{r}
exp(glmmodel$coefficients[2] + glmmodel$coefficients[3])
```
The odds are 3.800759 times greater.
