---
title: "Regression and Diagnostics"
author: "Brian Motherway"
output: github_document 
---
# Overview and Regression
In this project I am analyzing prices of houses in New Taipei City, Taiwan with that data set that can be found at this link: https://archive.ics.uci.edu/dataset/477/real+estate+valuation+data+set

```{r setup, message=FALSE, warning=FALSE, error=FALSE}
library(ISLR2)
library(car)
library(leaps)
library(msm)
library(tidyverse)
library(stringr)
library(GLMsData)
library(statmod)
library(MASS)
```

```{r}
#Loading my dataset
setwd("C:/Users/mothe/OneDrive/Documents/CMU/Mini 2/Statistical Foundations of Business Analytics/Data")
real_estate <- read_csv("real-estate-valuation-data-set.csv")
```
Fitting a multi linear regression and determining if there is any issue with collinearity by evaluating the Variance Inflation Factor (VIF).

```{r}
multi_reg <- lm(unit_price ~ convenience_stores + distance, data = real_estate)
multi_reg
vif(multi_reg)
```
Given both of the VIFs are below 5, I do not have concerns regarding collinearity of convenience_store and distance.

I plot each predictor against the unit_price to further analyze the relationship between each predictor and unit_price.

```{r}
reg_conv <- plot(unit_price ~ convenience_stores, data = real_estate)
reg_dis <- plot(unit_price ~ distance, data = real_estate)
```
Based off the plot distance and unit_price it looks like this relationship is not linear, but I want to do more investigating.

```{r}
#Evaluate all three linear models
multi_reg
conv_reg <- lm(unit_price ~ convenience_stores, data = real_estate)
dis_reg <- lm(unit_price ~ distance, data = real_estate)

#Utilize anove to conduct an F-Test comparison of the models 
anova(multi_reg, conv_reg, dis_reg, test = "F")

```
Based on the F-test for nested models, now we can conclude that using convenience_stores as the only predictor of unit price is preferable. Based on the plot of distance to unit price I can determine the relationship appears to be exponential. Here is the plot of the log(distance) to unit price. 

```{r}
log_plot_distance <- plot(unit_price ~ I(log(distance)), data = real_estate)
```
Relationship appears much more linear as suspected.

```{r}

#Plot multiple regression of unit_price against convenience_stores and log(distance)
multi_reg_log <- lm(unit_price ~ convenience_stores + I(log(distance)), data = real_estate)
multi_reg_log

#Evaluate which model is the best with an F-Test
anova(conv_reg, multi_reg, multi_reg_log, test = "F")
```
Model #3 has the lowest residual sum of squares, so this appears to be the best fit.

# Diagnostics
Now lets do some model diagnostics to confirm we have a good regression model.

```{r}
#Standardizing the residuals over the standard deviation of the residuals  
std_residuals <- rstandard(multi_reg)

#Plotting the standardized residuals of this model against the predictor convenience stores.
scatter.smooth(
  std_residuals ~ real_estate$convenience_stores,
  col = "grey",
  las = 1,
  ylab = "Standardized residuals",
  xlab = "Height (inches)"
)
res_conv_plot <- ggplot(data = real_estate, aes(x = real_estate$convenience_stores, y = std_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_conv_plot
```
Generally the residuals are fairly consistent with an average standardized residual of about 0.

```{r}
#Plotting the standardized residuals of this model against the predictor logarithm of distance.
res_logdis_plot <- ggplot(data = real_estate, aes(x = log(real_estate$distance), y = std_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_logdis_plot
```
The standardized residuals for value of log(distance) have a trend below zero.

```{r}
multi_reg_fitted <- fitted(multi_reg_log)
res_funitprice_plot <- ggplot(data = real_estate, aes(x = multi_reg_fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_funitprice_plot 
```
The standardized residuals for a unit price of greater than 50 is not averaging zero. We may be over estimating these points.

```{r}
qqnorm(std_residuals, las = 1, pch = 19)
qqline(std_residuals)
```
It looks like the model might be assuming tails that are slightly too heavy for the model.

# Box-Cox Method

Utilizing the Box-Cox method to find the optimal transformation of unit_price.

```{r}
box_cox <- boxcox(unit_price ~ convenience_stores + log(distance), data = real_estate)
lambda <- box_cox$x[which.max(box_cox$y)] 
lambda
```
Optimal lambda is found (0.42) - now we will run regression with that transformation and compare residuals.

```{r}
bc_mlr <- lm(((unit_price^lambda - 1)/ lambda)~ convenience_stores + log(distance),  data = real_estate)
summary(bc_mlr)
hist(multi_reg$residuals) #Original residuals
hist(bc_mlr$residuals) # Much more Normal-looking residuals.
```
Residuals appear much more normal under the Box-Cox method.

# Model Diagnostics of the Box-Cox Model
```{r}
#Standardized residuals
bcstd_residuals <- rstandard(bc_mlr)

#Plotting standarized residuals
scatter.smooth(
  bcstd_residuals ~ real_estate$convenience_stores,
  col = "grey",
  las = 1,
  ylab = "Standardized residuals",
  xlab = "Height (inches)"
)

bcres_conv_plot <- ggplot(data = real_estate, aes(x = real_estate$convenience_stores, y = bcstd_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_conv_plot
bcres_conv_plot
```
```{r}
bcres_logdis_plot <- ggplot(data = real_estate, aes(x = log(real_estate$distance), y = bcstd_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_logdis_plot
bcres_logdis_plot
```
```{r}
bc_mlr_fitted <- fitted(bc_mlr)
bcres_funitprice_plot <- ggplot(data = real_estate, aes(x = bc_mlr_fitted, y = bcstd_residuals)) +
  geom_point() +
  geom_smooth(se = FALSE)
res_funitprice_plot 
bcres_funitprice_plot
```
There is no material difference in the residuals for the boxcox plot.
```{r}
qqnorm(std_residuals, las = 1, pch = 19)
qqline(std_residuals)
qqnorm(bcstd_residuals, las = 1, pch = 19)
qqline(bcstd_residuals)
```
There is no material difference in the normality for the boxcox plot.

# Influential Observations

We will use DFBETA scores to determine if there are any concerning influential observations.
```{r}
bc_mlr_im <- influence.measures(bc_mlr)
summary(bc_mlr_im)
```
No observation for DFBETA were flagged as influential, we don't need to take any action to remove any observations.
